### 2017-06-09 Manuscripts/cog-psych/cogpsych.tex
after these changes, we are going to start WINNING again

Although we cannot know the precise colors that were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast. To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}. In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups was not used as a possible hue for the experimenter-defined categories). By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.

@@

Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation. Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known. However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with the psychological color boundaries perceived by participants. While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli. This precludes exemplar modeling of their data as it is necessary to encode exemplars in psychological space to determine the similarity between exemplars. By consequence, although these results likely indicate that contrast exerts {\em some} influence, they do not precisely describe the nature of that influence. In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.

@@

It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation. For example, it may be possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to include contrast, this mechanism would be at odds with the standard hierarchical Bayesian framework. In contrast, these ideas emerge naturally from the exemplar view: We have not modified any of the core elements of the GCM in defining the PACKER model, we simply weight exemplar similarity during aggregation. Thus, beyond formally specifying the role of contrast in creative generation, the PACKER account allows us to evaluate the well-understood principles of the exemplar view within the comparatively understudied field of conceptual generation. Nonetheless, we will discuss integrating our approaches in the General Discussion.

@@

To begin our investigation, we sought to extend the early evidence obtained from our analysis of the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials. We used an artificial stimulus design, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}. Unlike those used in the \cite{jern2013probabilistic} experiments, distance on the physical dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely. To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.

@@

As noted in the introduction of this paper, the task of generating members of a new category can be aligned with common tasks studied in the categorization literature: Whereas classification consists of predicting an object's category label on the basis of its features, inference consists of predicting an observed feature, given a set of observed features and a category label. Generation thus consists of predicting {\em all} features of an object, given a novel category label. In keeping with this insight, we designed our generation task as an extension of the traditional artificial classification learning paradigm. The task differs from traditional work in creative cognition primarily through the use of an artificial domain, which enables the application of computational models. Unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).

@@

Thus, the core results of Experiment 1 indicate that generated categories can systematically differ from what we would expect based on prior work. The negative (or zero) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties. That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure. In Experiment 2, we test this claim more systematically.

@@

Table \ref{table:global-model-fits} contains the results of this fitting procedure. Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters). The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression})

@@

\begin{figure}[p]
    \begin{center}
    \inputpgf{figs/}{range-diff-gradients.pgf}
    \caption{Behavioral and simulated range difference gradients. Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' (orange) versus horizontally aligned `rows' (purple).}
    \nbcnote{Need to run this on lando with many more samples. current = 30}
    \label{fig:range-diff-gradients}
    \end{center}
\end{figure}

@@

Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms that explain them. Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: The broad and strong influence of category contrast, and the interdependence between category location and distributional structure. It may be possible to combine the hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure. However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of a traditional, semi-conjugate Bayesian approach. This suggests that category generation is a fundamentally different computational-level problem \citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}. Characterizing that problem and conducting a rational analysis is an important direction for future research. To that aim, we plan to explore the connection between exemplar modeling as an Importance Sampling approximation \citep{shi10exemplar}, and see what sort of computational-level problem PACKER approximates. Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model. Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models, such as SUSTAIN \citep{lovemg04} or DIVA \citep{}.

@@

In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a semi-conjugate model. Their model acts very similar to our version of it and receives comparable fits.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$). Category covariance matrices (specifying variance and covariance along $k$-dimensions, are assumed to be Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$ ($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all categories are viewed as samples. Assuming a given $\Sigma_D$, a category covariance matrix $\Sigma$ can be computed on the basis of its examples:

>>>>

Although we cannot know the precise colors that were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast. To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}. In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups were not used as a possible hue for the experimenter-defined categories). By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.

@@

Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation. Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known. However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with the psychological color boundaries perceived by participants. While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli. This precludes traditional applications of categorization models to their data as it is usually necessary to encode objects in psychological space in order to accurately determine the similarity between objects. By consequence, although these results likely indicate that contrast exerts {\em some} influence, they do not precisely describe the nature of that influence. In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.

@@

It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation. For example, it may be possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to include contrast, but this mechanism would be at odds with the standard hierarchical Bayesian framework. In contrast, these ideas emerge naturally from the exemplar view: We have not modified any of the core elements of the GCM in defining the PACKER model, we simply weight exemplar similarity during aggregation. Thus, beyond formally specifying the role of contrast in creative generation, the PACKER account allows us to evaluate the well-understood principles of the exemplar view within the comparatively understudied field of conceptual generation. Nonetheless, we will discuss integrating our approaches in the General Discussion.

@@

To begin our investigation, we sought to extend the early evidence obtained from our analysis of the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials. We used an artificial stimulus design: A two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}. Unlike those used in the \cite{jern2013probabilistic} experiments, distance on the physical dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely. To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.

@@

As noted in the introduction of this paper, the task of generating members of a new category can be aligned with common tasks studied in the categorization literature: Whereas classification consists of predicting an object's category label on the basis of its features, inference consists of predicting an observed feature, given a set of observed features and a category label. Generation thus consists of predicting {\em all} features of an object, given a novel category label. In keeping with this insight, we designed our generation task as an extension of the traditional artificial classification learning paradigm. The task differs from traditional work in creative cognition primarily through the use of an artificial domain, which enables the application of computational models. The use of an artificial domain also requires the addition of a training phase, during which participants learn about the categories in the domain. As a result,, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).

@@

Thus, the core results of Experiment 1 indicate that generated categories can systematically differ from what we would expect based on prior work. The negative (or null) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties. That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure. In Experiment 2, we test this claim more systematically.

@@

Table \ref{table:global-model-fits} contains the results of this fitting procedure. Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters). The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression}).

@@

\begin{figure}[p]
    \begin{center}
    \inputpgf{figs/}{range-diff-gradients.pgf}
    \caption{Behavioral and simulated range difference gradients. Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' (orange) versus horizontally aligned `rows' (purple).}
    \label{fig:range-diff-gradients}
    \end{center}
\end{figure}

@@

Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms that explain them. Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: The broad and strong influence of category contrast, and the interdependence between category location and distributional structure. It may be possible to combine the hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure. However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of a traditional, semi-conjugate Bayesian approach. This suggests that category generation is a fundamentally different computational-level problem \citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}. Characterizing that problem and conducting a rational analysis is an important direction for future research. To that aim, we plan to explore the connection between exemplar modeling as an Importance Sampling approximation \citep{shi10exemplar}, and see what sort of computational-level problem PACKER approximates. Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model. Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2015human,smith2000thirty}.

@@

In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a semi-conjugate model. Their model acts very similarly to our version.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$). Category covariance matrices (specifying variance and covariance along $k$-dimensions), are assumed to be Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$ ($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all categories are viewed as samples. Assuming a given $\Sigma_D$, a category covariance matrix $\Sigma$ can be computed on the basis of its examples:
